---
title: "Rosette Damage x Mowing Data Analysis"
author: "Emma van der Heide"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data Analysis for Rosette Damage x Mowing Analysis

```{r load data and packages}
#set working directory
setwd("~/Work that isn't on OneDrive/Rosette_Damage_x_Mowing")

#load data
dat <- read.csv("ManuscriptData1.csv")

#make data into factors
dat$Block <- as.factor(dat$Block)
dat$SubBlock <- as.factor(dat$SubBlock)
dat$Treatment <- as.factor(dat$Treatment)
dat$RosDam <- as.factor(dat$RosDam)
dat$EarlyMow <- as.factor(dat$EarlyMow)
dat$LateMow <- as.factor(dat$LateMow)

#load required packages
library(lme4)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(ggplot2)
library(dplyr)
library(multcomp)
```

#Flowering Date

Analysis of date of first flowering (days after May 2nd), defined as the date on which each individual plant had its first anthesed capitulum, after that plant had received all physical damage treatments.

```{r}
#FIT WITH TREATMENT BUT ALSO TRY TO MAKE IT WORK WITH Y/N OF RD/EM/LM
#fit model and summarize
#add LLL at a later date
fitFD <- glmmTMB(FloweringDate_doe ~ RosDam*EarlyMow*LateMow + LLLMay2 + (1|Block), data = dat, family = "gaussian")
summary(fitFD)

#check residuals
sim.fitFD<-simulateResiduals(fitFD)
plot(sim.fitFD) #visual inspection of residuals: looks good

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmFD <- emmeans(fitFD, specs=~RosDam*EarlyMow*LateMow, type="response")
emmFD

#get compact letter display for pairwise comparisons
cld(emmFD, Letters=letters)

#Visualize output
#Put emmeans output into a data frame
FlDate<-as.data.frame(cld(emmFD, Letters=letters))
#Add Treatment Name to Data Frame
FlDate<-mutate(FlDate, Treatment=case_when(
  RosDam == "n" & EarlyMow == "n" & LateMow == "n" ~ "Control",
  RosDam == "y" & EarlyMow == "n" & LateMow == "n" ~ "RD",
  RosDam == "y" & EarlyMow == "y" & LateMow == "n" ~ "RD+EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "n" ~ "EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "y" ~ "EM+LM",
  RosDam == "n" & EarlyMow == "n" & LateMow == "y" ~ "LM",
  RosDam == "y" & EarlyMow == "y" & LateMow == "y" ~ "RD+EM+LM",
  RosDam == "y" & EarlyMow == "n" & LateMow == "y" ~ "RD+LM",
))
#Scale data smaller for visualization (currently not using this in chart)
FlDate$mean<-FlDate$emmean/10
FlDate$LCL<-FlDate$lower.CL/10
FlDate$UCL<-FlDate$upper.CL/10

#make a plot
FlDatePlot<-ggplot(FlDate,aes(y=emmean, x=Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Flowering Date (Day of Experiment)")+
    geom_text(aes(label = .group, y = upper.CL + 5), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ylim(0,90)+
  ggtitle("Average date of first flowering")+
  geom_errorbar(aes(x=Treatment, ymin=lower.CL, ymax=upper.CL))
FlDatePlot
```

#Height at Flowering
```{r Height at Flowering}
#fit model and summarize results
fitFlHt <- glmmTMB(HeightatFlowering ~ RosDam*EarlyMow*LateMow + LLLMay2+ (1|Block), data = dat, family = "gaussian")
summary(fitFlHt)

#check residuals
sim.fitFlHt<-simulateResiduals(fitFlHt)
plot(sim.fitFlHt) #visual inspection: looks great

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmFlHt <- emmeans(fitFlHt, specs=~RosDam*EarlyMow*LateMow, type="response")
emmFlHt

#get compact letter display for pairwise comparisons
cld(emmFlHt, Letters=letters)

#Visualize results
#Put emmeans output into a data frame
FlHt<-as.data.frame(cld(emmFlHt, Letters=letters))
#Add Treatment Name to Data Frame
FlHt<-mutate(FlHt, Treatment=case_when(
  RosDam == "n" & EarlyMow == "n" & LateMow == "n" ~ "Control",
  RosDam == "y" & EarlyMow == "n" & LateMow == "n" ~ "RD",
  RosDam == "y" & EarlyMow == "y" & LateMow == "n" ~ "RD+EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "n" ~ "EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "y" ~ "EM+LM",
  RosDam == "n" & EarlyMow == "n" & LateMow == "y" ~ "LM",
  RosDam == "y" & EarlyMow == "y" & LateMow == "y" ~ "RD+EM+LM",
  RosDam == "y" & EarlyMow == "n" & LateMow == "y" ~ "RD+LM",
))

#make a plot
FlHtPlot<-ggplot(FlHt,aes(y=emmean, x=Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Height at Flowering")+
    geom_text(aes(label = .group, y = upper.CL + 10), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  ylim(0, 130)+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ggtitle("Height at Flowering")+
  geom_errorbar(aes(x=Treatment, ymin=lower.CL, ymax=upper.CL))
FlHtPlot
```

#Analyze Maximum Number of Capitula

Maximum number of capitula is the max number of capitula + buds that were ever present on a plant, after that individual plant had had all physeical damage treatments. This is because some capitula drop off/blow away as they senesce, meaning that the final assessment of number of capitula may be inaccurate.

```{r Max. Cap.}
#fit model and summarize
fitMC <- glmmTMB(MaxCapitula ~ RosDam*EarlyMow*LateMow + LLLMay2 + (1|Block), data = dat, family = "poisson") #poisson error distribution chosen because this is count data
summary(fitMC)

#check assumptions by examining residuals
sim.fitMC <- simulateResiduals(fitMC)
plot(sim.fitMC) #Visual inspection: this looks fine, dispersion test not concerning

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmMC <- emmeans(fitMC, specs=~RosDam*EarlyMow*LateMow, type="response")
emmMC

#get compact letter display for pairwise comparisons
cld(emmMC, Letters=letters)

#Visualize results
#Put emmeans output into a data frame
MaxCap<-as.data.frame(cld(emmMC, Letters=letters))
#Add Treatment Name to Data Frame
MaxCap<-mutate(MaxCap, Treatment=case_when(
  RosDam == "n" & EarlyMow == "n" & LateMow == "n" ~ "Control",
  RosDam == "y" & EarlyMow == "n" & LateMow == "n" ~ "RD",
  RosDam == "y" & EarlyMow == "y" & LateMow == "n" ~ "RD+EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "n" ~ "EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "y" ~ "EM+LM",
  RosDam == "n" & EarlyMow == "n" & LateMow == "y" ~ "LM",
  RosDam == "y" & EarlyMow == "y" & LateMow == "y" ~ "RD+EM+LM",
  RosDam == "y" & EarlyMow == "n" & LateMow == "y" ~ "RD+LM",
))

#make a plot
MaxCapPlot<-ggplot(MaxCap,aes(y=rate, Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Capitula Production")+
  geom_text(aes(label = .group, y = asymp.UCL + 1), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  ylim(0, 17)+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ggtitle("Maximum Number of Capitula Present on Plant, Post-Treatment")+
  geom_errorbar(aes(x=Treatment, ymin=asymp.LCL, ymax=asymp.UCL))
MaxCapPlot
```

#Analyze Maximum Number of Capitula, including those capitula that successfully produced seeds before all treatments were complete.

Same as above, but we dissected removed plant matter (some late mowed plants produced capitula before the late mow), and added the number that produced viable seeds (visually determined) to the maximum capitula numbers from the previous fit

```{r Max. Cap.}
#fit model and summarize
fitMCplus <- glmmTMB(MaxCapitulaIncPreTrtSeedProducers ~ RosDam*EarlyMow*LateMow + LLLMay2 + (1|Block), data = dat, family = "poisson") #poisson error distribution chosen because this is count data
summary(fitMCplus)

#check assumptions by examining residuals
sim.fitMCplus <- simulateResiduals(fitMCplus)
plot(sim.fitMCplus) #Visual inspection: this looks good

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmMCplus <- emmeans(fitMCplus, specs=~RosDam*EarlyMow*LateMow, type="response")
emmMCplus

#get compact letter display for pairwise comparisons
cld(emmMCplus, Letters=letters)

#Visualize results
#Put emmeans output into a data frame
MaxCapPlus<-as.data.frame(cld(emmMCplus, Letters=letters))
#Add Treatment Name to Data Frame
MaxCapPlus<-mutate(MaxCapPlus, Treatment=case_when(
  RosDam == "n" & EarlyMow == "n" & LateMow == "n" ~ "Control",
  RosDam == "y" & EarlyMow == "n" & LateMow == "n" ~ "RD",
  RosDam == "y" & EarlyMow == "y" & LateMow == "n" ~ "RD+EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "n" ~ "EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "y" ~ "EM+LM",
  RosDam == "n" & EarlyMow == "n" & LateMow == "y" ~ "LM",
  RosDam == "y" & EarlyMow == "y" & LateMow == "y" ~ "RD+EM+LM",
  RosDam == "y" & EarlyMow == "n" & LateMow == "y" ~ "RD+LM",
))

#make a plot
MaxCapPlusPlot<-ggplot(MaxCapPlus,aes(y=rate, Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Capitula Production")+
  geom_text(aes(label = .group, y = asymp.UCL + 1), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  ylim(0, 17)+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ggtitle("Maximum Number of Capitula Present on Plant, Including Pre-Treatment Seed Producers")+
  geom_errorbar(aes(x=Treatment, ymin=asymp.LCL, ymax=asymp.UCL))
MaxCapPlusPlot
```

#Maximum Height

This is maximum post-treatment height (after all physical damages had been applied)

```{r}
#fit model and summarize
fitMaxHt<-glmmTMB(MaxHeight~RosDam*EarlyMow*LateMow+LLLMay2+(1|Block), data=dat, family="nbinom1")
summary(fitMaxHt)

#check residuals
sim.fitMaxHt<-simulateResiduals(fitMaxHt)
plot(sim.fitMaxHt) #looks fine

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmMaxHt <- emmeans(fitMaxHt, specs=~RosDam*EarlyMow*LateMow, type="response")
emmMaxHt

#get compact letter display for pairwise comparisons
cld(emmMaxHt, Letters=letters)

#Visualize
#Put emmeans output into a data frame
MaxHt<-as.data.frame(cld(emmMaxHt, Letters=letters))
#Add Treatment Name to Data Frame
MaxHt<-mutate(MaxHt, Treatment=case_when(
  RosDam == "n" & EarlyMow == "n" & LateMow == "n" ~ "Control",
  RosDam == "y" & EarlyMow == "n" & LateMow == "n" ~ "RD",
  RosDam == "y" & EarlyMow == "y" & LateMow == "n" ~ "RD+EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "n" ~ "EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "y" ~ "EM+LM",
  RosDam == "n" & EarlyMow == "n" & LateMow == "y" ~ "LM",
  RosDam == "y" & EarlyMow == "y" & LateMow == "y" ~ "RD+EM+LM",
  RosDam == "y" & EarlyMow == "n" & LateMow == "y" ~ "RD+LM",
))

#make a plot
MaxHtPlot<-ggplot(MaxHt,aes(y=response, x=Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Maximum Height (cm)")+
  geom_text(aes(label = .group, y = asymp.UCL + 10), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ylim(0,160)+
  ggtitle("Maximum Height")+
  geom_errorbar(aes(x=Treatment, ymin=asymp.LCL, ymax=asymp.UCL))
MaxHtPlot
```

#Number of Stems
```{r}
#fit model and summarize
fitNS<-glmmTMB(NoStemsFinal~RosDam*EarlyMow*LateMow+NoStemsMay2+(1|Block), data=dat, family="poisson")
summary(fitNS)


#check residuals
sim.fitNS<-simulateResiduals(fitNS)
plot(sim.fitNS) #looks fine

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmNS <- emmeans(fitNS, specs=~RosDam*EarlyMow*LateMow, type="response")
emmNS

#get compact letter display for pairwise comparisons
cld(emmNS, Letters=letters)

#Visualize
#Put emmeans output into a data frame
NoStems<-as.data.frame(cld(emmNS, Letters=letters))
#Add Treatment Name to Data Frame
NoStems<-mutate(NoStems, Treatment=case_when(
  RosDam == "n" & EarlyMow == "n" & LateMow == "n" ~ "Control",
  RosDam == "y" & EarlyMow == "n" & LateMow == "n" ~ "RD",
  RosDam == "y" & EarlyMow == "y" & LateMow == "n" ~ "RD+EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "n" ~ "EM",
  RosDam == "n" & EarlyMow == "y" & LateMow == "y" ~ "EM+LM",
  RosDam == "n" & EarlyMow == "n" & LateMow == "y" ~ "LM",
  RosDam == "y" & EarlyMow == "y" & LateMow == "y" ~ "RD+EM+LM",
  RosDam == "y" & EarlyMow == "n" & LateMow == "y" ~ "RD+LM",
))

#make a plot
NoStemsPlot<-ggplot(NoStems,aes(y=rate, x=Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Number of Stems")+
  geom_text(aes(label = .group, y = asymp.UCL + .5), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ylim(0,5)+
  ggtitle("Number of Stems at 10 cm, at Experiment Termination")+
  geom_errorbar(aes(x=Treatment, ymin=asymp.LCL, ymax=asymp.UCL))
NoStemsPlot
```

