---
title: "Rosette Damage x Mowing Data Analysis"
author: "Emma van der Heide"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data Analysis for Rosette Damage x Mowing Analysis

```{r load data and packages}
#set working directory
setwd("~/Work that isn't on OneDrive/Rosette_Damage_x_Mowing")

#load data
dat <- read.csv("ManuscriptData1.csv")

#make data into factors
dat$Block <- as.factor(dat$Block)
dat$SubBlock <- as.factor(dat$SubBlock)
dat$Treatment <- as.factor(dat$Treatment)
dat$RosDam <- as.factor(dat$RosDam)
dat$EarlyMow <- as.factor(dat$EarlyMow)
dat$LateMow <- as.factor(dat$LateMow)
dat$SurviveJuly5 <- as.factor(dat$SurviveJuly5)

#load required packages
library(lme4)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(ggplot2)
library(dplyr)
library(multcomp)
```

#Analyze Maximum Number of Capitula

Maximum number of capitula is the max number of capitula + buds that were ever present on a plant, after the last mowing treatment happened on (WHICH DATE?)
This is because some capitula drop off/blow away as they senesce, meaning that the final assessment of number of capitula may be inaccurate.

```{r Max. Cap.}
#fit model and summarize
fit1 <- glmmTMB(MaxCapitula ~ Treatment + LLLMay2 + (1|Block), data = dat, family = "poisson") #poisson error distribution chosen because this is count data
summary(fit1)

#check assumptions by examining residuals
sim.fit1 <- simulateResiduals(fit1)
plot(sim.fit1) #Visual inspection: this looks great

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmMaxCap <- emmeans(fit1, specs=pairwise~Treatment, type="response")
emmMaxCap

#get compact letter display for pairwise comparisons
cld(emmMaxCap, Letters=letters)

#Visualize results
#Put emmeans output into a data frame
MaxCap<-as.data.frame(cld(emmMaxCap, Letters=letters))
#make a plot
MaxCapPlot<-ggplot(MaxCap,aes(y=rate, Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Capitula Production")+
  geom_text(aes(label = .group, y = max(rate) + 3.5), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  ylim(0, 14)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ggtitle("Maximum Number of Capitula Present on Plant")+
  geom_errorbar(aes(x=Treatment, ymin=asymp.LCL, ymax=asymp.UCL))
MaxCapPlot
```

#Maximum capitula conclusions:

#Height at Flowering
```{r Height at Flowering}
#fit model and summarize results
fit2 <- glmmTMB(HeightatFlowering ~ Treatment + (1|Block) + LLLMay2, data = dat, family = "gaussian")
summary(fit2)

#check residuals
sim.fit2<-simulateResiduals(fit2)
plot(sim.fit2) #visual inspection: looks great

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmHtFl <- emmeans(fit2, specs=pairwise~Treatment, type="response")
emmHtFl

#get compact letter display for pairwise comparisons
cld(emmHtFl, Letters=letters)

#Visualize results
#Put emmeans output into a data frame
HtFl<-as.data.frame(cld(emmHtFl, Letters=letters))
#make a plot
HtFlPlot<-ggplot(HtFl,aes(y=emmean, x=Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Height at Flowering")+
    geom_text(aes(label = .group, y = max(emmean) + 20), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  ylim(0, 130)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ggtitle("Height at Flowering")+
  geom_errorbar(aes(x=Treatment, ymin=lower.CL, ymax=upper.CL))
HtFlPlot
```

#Height at Flowering Conclusions:

#Survival

Analysis of survival is in a separate document. Below is some code I was playing with initially that wasn't working the way I intended. Will eventually delete this part from this document.
In future, could look at how survivorship decreases through time (the full curve). See Trevor's seed dispersal paper, code is in Emma's downloads folder.

#Survival analysis, based on Trevor's paper

```{r}
#Trevor's functions for calculating means and ks.test are moved into Survival Analysis.R; I don't think I need to follow that through for this analysis


##### Convert data to format suitable for survival model functions ----------------------------------------

# Construct concise representation of original dataset
# ToD indicates time of death
# Cens indicates censor status (1 = dead, 0 = censored and survived until end)
for(i in 1:nrow(surv_dat)){ #initialize the main data frame, iterates over every row of "surv_dat" ("surv_dat" formerly "Data")
  if(i == 1){
    df_main <- matrix(ncol = 7, nrow = 0)} #creates an empty matrix with 7 columns - NOT SURE WHY 7?????
  row_sub <- surv_dat[i, ] #The current row from Data is stored in row_sub
  for(j in 7:ncol(surv_dat)){ #nested loop that goes through columns starting with column 7 - SHOULD MAYBE BE 6 BECAUSE SURVIVAL DATA STARTS IN COLUMN 6? OR 5 IF WE'RE DOING -BLOCK
    if(j == 7){
      prevNum <- 1 #where the data starts/maximum number of seeds, formerly 25
      df_sub <- matrix(ncol = 7, nrow = 0)} #holds temporary results
    curNum <- as.numeric(row_sub[j])
    if(curNum < prevNum){ #If the current number (from row_sub) is less than prevNum, it creates new rows for df_sub
      df_sub <- rbind(df_sub, matrix(c(as.character(row_sub[1:5])), #str_remove(names(Data)[j], "t_"), 1), #It uses rbind() to add rows indicating that the subject died (Cens = 1) at a particular time point. The str_remove() function removes the prefix t_ from the column names to get a more readable time identifier.
                                     nrow = prevNum - curNum, ncol = 7, byrow = TRUE))}
    if(j == ncol(surv_dat) & curNum > 0){
      df_sub <- rbind(df_sub, matrix(c(as.character(row_sub[1:5])), #str_remove(names(Data)[j], "t_"), 0),
                                     nrow = curNum, ncol = 7, byrow = TRUE))}
    prevNum <- curNum} #At the last column, if the current number is greater than 0, it adds rows indicating the subject survived until the end of the study (Cens = 0). prevNum is updated to the current number for the next iteration.
  df_main <- rbind(df_main, df_sub)} #results stored in df_sub are combined into df_main
#This gives a big error: "Warning: data length [5] is not a sub-multiple or multiple of the number of columns [7]" FIGURE OUT WHAT THIS MEANS AND FIX IT!!!!


df_main <- data.frame(df_main, stringsAsFactors = FALSE) #df_main is converted from a matrix to a data frame.
names(df_main) <- c(names(surv_dat)[1:5], "ToD", "Cens") #column names are set to the first five column names of the original dataset, followed by ToD and Cens. NOTE THAT THIS INCLUDES MAY.2 FOR SOME REASON
df_main$Row <- as.numeric(df_main$Row) #convert columns to numeric
df_main$Group <- as.numeric(df_main$Group)
df_main$Plant <- as.numeric(df_main$Plant)
df_main$Treatment <- as.numeric(df_main$Treatment)
df_main$ToD <- as.numeric(df_main$ToD)
df_main$Cens <- as.numeric(df_main$Cens)
DataAlt <- df_main #final data frame is stored in DataAlt

# Remove temporary variables since they will no longer be used
remove(df_main, df_sub, row_sub, curNum, prevNum, i, j)

#From Trevor's "stats" script
# Total proportion of seeds removed after a given time - we could get proportion of plants dead after a given time, can also get proportion by treatment by using one of the 8 treatment-specific datasets made above
1 - mean(surv_dat$May.2); sd(surv_dat$May.2)/sqrt(length(surv_dat$May.2))    # May 2nd - 0 means all values are 1, average is 1 meaning survival because of subtraction
1 - mean(surv_dat$May.30); sd(surv_dat$May.30)/sqrt(length(surv_dat$May.30))    # May 30th
1 - mean(surv_dat$Jul.25); sd(surv_dat$Jul.25)/sqrt(length(surv_dat$Jul.25))    # July 25th

#GLMS for seed removal (or plant death over time) using proportion data - see Trevor's code. I don't think we need to do this because we don't have replicated proportion data
GLM_Jul25<- glmer(cbind(1 - surv_dat$Jul.25, surv_dat$Jul.25) ~ Treatment + (1|Block), data = surv_dat, family = "binomial")
summary(GLM_Jul25)
#THIS DOES NOT WORK FOR ME - I THINK BECAUSE I'M NOT STARTING WITH 25 SEEDS; RATHER A SINGLE PLANT

#Not sure we need to do this next part
##### Qualitative comparison of GLMs and survival regressions ---------------------------------------------

# Some sort of accelerated failure model (increasing hazard over time) makes most sense
# Will try exponential and Weibull

# Note: preferred method frailty() for random effect in survival model does NOT work in survreg
# Encoding random effect as (1|Block) also does not work; will thus encode as fixed effect
# This is perfectly valid, just does not let us make numeric predictions outside context of experiment

# CA, exponential hazard
Surv1_CA <- survreg(Surv(ToD, Cens) ~ Warmed + Elaiosome + Warmed:Elaiosome + as.factor(Block),
                    data = DataAlt_CA, dist = "exponential")
summary(Surv1_CA)
AIC(Surv1_CA)

# CA, Weibull (default) hazard
Surv2_CA <- survreg(Surv(ToD, Cens) ~ Warmed + Elaiosome + Warmed:Elaiosome + as.factor(Block),
                    data = DataAlt_CA)
summary(Surv2_CA)
AIC(Surv2_CA)

#Plots, based on Trevor's code, more info in that script including for legends, etc.

#Compare two treatments - we could maybe do this with all 8?
# CN Unwarmed: E+ (dark green) v E- (light green)
print(surv.plots(Data_CN_NW_YE, Data_CN_NW_NE, PlotColours[1], PlotColours[2],
                 bottom = FALSE, left = TRUE, atext = "CN Unwarmed"),
      vp = viewport(layout.pos.row = 25:500, layout.pos.col = 25:700)) #this last line might not be necessary

#first, define function to plot 2 survival curves
surv.plots <- function(df1, df2, colour1, colour2, bottom, left, atext){
  
  # Get p-value from K-S test
  pval <- time.ks(df1, df2)
  pval <- ifelse(pval < 0.001, "italic(p) < 0.001",
                 paste("paste(italic(p), \" =\")",
                       paste0("\"", sprintf("%0.3f", round(pval, 3)), "\""), sep = "~"))
  
  # Plot survival curves
  ggplot() +
    geom_point(data = time.means(df1), aes(x = Time, y = Mean), colour = colour1, size = 0.01) +
    geom_line(data = time.means(df1), aes(x = Time, y = Mean), colour = colour1, size = 0.25) +
    geom_errorbar(data = time.means(df1), aes(x = Time, ymin = Mean - SEM, ymax = Mean + SEM),
                  width = 0.3, colour = colour1, size = 0.2) +
    geom_point(data = time.means(df2), aes(x = Time, y = Mean), colour = colour2, size = 0.01) +
    geom_line(data = time.means(df2), aes(x = Time, y = Mean), colour = colour2, size = 0.25) +
    geom_errorbar(data = time.means(df2), aes(x = Time, ymin = Mean - SEM, ymax = Mean + SEM),
                  width = 0.3, colour = colour2, size = 0.2) +
    coord_cartesian(ylim = c(0, 1)) +
    scale_x_continuous(expand = c(0.01, 0.01), limits = c(0, 15), 
                       breaks = c(seq(1,14, by = 1)),
                       labels = c(0, rep("", 21))) +
    annotate("text", x = 47.7, y = 25, label = atext, hjust = 1, size = 1.3) +
    annotate("text", x = 47.7, y = 23.4, label = pval, hjust = 1, size = 1.3, parse = TRUE) +
    xlab("Time (Date)") +
    ylab("Plants Remaining") +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(colour = "gray90", size = 0.2),
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.border = element_rect(colour = "black", fill = NA, size = 0.4),
          panel.background = element_rect(fill = "white"),
          axis.text.x = element_text(size = 3.5),
          axis.text.y = element_text(size = 3.5),
          axis.title.x = element_text(size = 4),
          axis.title.y = element_text(size = 4),
          axis.ticks = element_line(colour = "black", size = 0.2),
          axis.ticks.length = unit(0.04, "cm")) -> graph
  
  # Format graphs depending on panel placement
  if(bottom == FALSE & left == TRUE){
    graph + theme(axis.ticks.x = element_blank(),
                  axis.text.x = element_blank(),
                  axis.title.x = element_blank(),
                  plot.margin = unit(c(0.08, 0.03, 0.43, 0.06), "cm"))} -> graph
  if(bottom == TRUE & left == TRUE){
    graph + theme(plot.margin = unit(c(0.01, 0.03, 0.08, 0.06), "cm"))} -> graph
  if(bottom == FALSE & left == FALSE){
    graph + theme(axis.ticks = element_blank(),
                  axis.text.x = element_blank(),
                  axis.text.y = element_blank(),
                  axis.title.x = element_blank(),
                  axis.title.y = element_blank(),
                  plot.margin = unit(c(0.08, 0.10, 0.43, 0.42), "cm"))} -> graph
  if(bottom == TRUE & left == FALSE){
    graph + theme(axis.ticks.y = element_blank(),
                  axis.text.y = element_blank(),
                  axis.title.y = element_blank(),
                  plot.margin = unit(c(0.01, 0.10, 0.08, 0.42), "cm"))} -> graph
  
  # Output graph
  return(graph)}

print(surv.plots(surv_dat_control, surv_dat_RDEMLM, PlotColours[1], PlotColours[2], bottom = FALSE, left = TRUE, atext="Control vs. All Damage"))

# Plot survival curves
print(surv.plots4(Data_CN_NW_NE, Data_CN_NW_YE, Data_CN_YW_NE, Data_CN_YW_YE,
                  PlotColours_All[1], PlotColours_All[3], PlotColours_All[5], PlotColours_All[7], TRUE),
      vp = viewport(layout.pos.row = 25:775, layout.pos.col = 25:1375))


```


#Flowering Date

Analysis of date of first flowering. 

```{r}
#fit model and summarize
#add LLL at a later date
fit4 <- glmmTMB(FloweringDate_doe ~ Treatment + (1|Block), data = dat, family = "gaussian")
summary(fit4)

#check residuals
sim.fit4<-simulateResiduals(fit4)
plot(sim.fit4) #visual inspection of residuals: looks good

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmFlDate <- emmeans(fit4, specs=pairwise~Treatment, type="response")
emmFlDate

#get compact letter display for pairwise comparisons
cld(emmFlDate, Letters=letters)

#Visualize output
#Put emmeans output into a data frame
FlDate<-as.data.frame(cld(emmFlDate, Letters=letters))
#Scale data smaller for visualization
FlDate$mean<-FlDate$emmeans.emmean/10
FlDate$LCL<-FlDate$emmeans.lower.CL/10
FlDate$UCL<-FlDate$emmeans.upper.CL/10

#make a plot
FlDatePlot<-ggplot(FlDate,aes(y=emmean, x=Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Flowering Date (Day of Experiment)")+
    geom_text(aes(label = .group, y = max(emmean) + 10), #+ number specifies how high the cld appears
            position = position_dodge(width = 0.75))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ylim(0,90)+
  ggtitle("Average date of first flowering")+
  geom_errorbar(aes(x=Treatment, ymin=lower.CL, ymax=upper.CL))
FlDatePlot
```

#Maximum Height (after mowing is complete)

```{r}
#fit model and summarize
fit5<-glmmTMB(MaxHeight~Treatment+LLLMay2+(1|Block), data=dat, family="nbinom1")
summary(fit5)

#check residuals
sim.fit5<-simulateResiduals(fit5)
plot(sim.fit5)

#calculate treatment means (estimated marginal means) and pairwise comparisons by treatment
emmMaxHt <- emmeans(fit5, specs=pairwise~Treatment, type="response")
emmMaxHt

#Visualize
#Put emmeans output into a data frame
MaxHt<-as.data.frame(emmMaxHt[1])

#make a plot
MaxHtPlot<-ggplot(MaxHt,aes(y=emmeans.response, x=emmeans.Treatment))+
  geom_bar(stat = "identity")+
  theme(panel.background = element_blank(), panel.border = element_rect(color = "black",fill=NA, size=1))+
  labs(x="Treatment", y="Maximum Height")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #angle text 45 degrees
  ylim(0,150)+
  ggtitle("Maximum Height (cm)")+
  geom_errorbar(aes(x=emmeans.Treatment, ymin=emmeans.asymp.LCL, ymax=emmeans.asymp.UCL))
MaxHtPlot
```

